---
title: "Data Preprocessing"
author: "Michael Lindner, Antonia Schuster"
date: "11/19/2021"
bibliography: bibliography.bib
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# When knitting the document the root dir has to be set
# knitr::opts_knit$set(root.dir = "..")
# This is now done via Rstudio global options.
```

# Preliminaries: Version control

For exact reproducibility of our results we use the package manager `renv` to keep track of the versions of all packages used.

To reproduce our analysis, install `renv` then use `renv::activate()` to activate the environment specified in the lockfile. You might have to use `renv::hydrate()` to install required packages as well. Afterwards you can run the R code as you would normally do.

# Preprocessing

## Data loading and variable selection


### Survey data

```{r message=FALSE}
library("haven")
i_hhresp <- read_sav("data/raw/i_hhresp.sav")

# Select relevant fields
library("dplyr")
df1 <- dplyr::select(i_hhresp, i_hidp, i_psu, i_strata, i_hhdenui_xw, i_gor_dv,
                     i_intdatey, i_fuelhave1, i_fuelhave2, i_fuelhave3, 
                     i_fuelhave4, i_fuelhave96, i_fuelduel, i_xpduely, 
                     i_duelpay, i_xpelecy,i_elecpay, i_xpgasy, i_gaspay, 
                     i_xpoily, i_xpsfly, i_hhdenui_xw, i_urban_dv, i_hhsize, 
                     i_fihhmnnet1_dv, i_nemp_dv, i_nue_dv, i_hhtype_dv, 
                     i_hsownd, i_ff_hsbeds, i_hhtype_dv, i_fborn, i_nkids_dv, 
                     i_xphsdb, i_ieqmoecd_dv, i_hsbeds, i_hsrooms, i_npensioner,
                     i_nadoecd_dv, i_nchoecd_dv, i_ncouple_dv, i_nonepar_dv,
                     i_heatch)


# remove large dataframe
rm(i_hhresp)
```

We will use equivalised household net income according to the OECD scale for our
analysis. Here we add the corresponding column as outlined in the USoc manual.

```{r}
df1$eq_income <- (df1$i_fihhmnnet1_dv / df1$i_ieqmoecd_dv)
```



# Summarising individual level information at the household level

## Data loading
```{r results=FALSE}
dfi <- read_sav("data/raw/i_indresp.sav")
dfind <- dplyr::select(dfi, i_hidp, pidp, i_health, i_sex_dv, i_aidhh,
                       i_hiqual_dv, i_age_dv)
rm(dfi)
```

## People with health condition
```{r}

dfind$flag_hcond <- dfind$i_health == 1 

dfind_nhealth <- dfind %>%
                 group_by(i_hidp) %>%
                 count(flag_hcond, name="nhcond") %>%
                 filter(flag_hcond) %>%
                 select(i_hidp, nhcond) %>%
                 full_join(dfind, by="i_hidp") %>%
                 select(i_hidp, nhcond) %>%
                 distinct(i_hidp, nhcond) %>%
                 mutate_at(vars(nhcond), ~replace(., is.na(.), 0))
            
```
### Merge data frames

Make sure to execute this only once
```{r}
df1$from_hh <- 1
dfind_nhealth$from_indall <- 1

df1 <- full_join(dfind_nhealth, df1, by="i_hidp")


```

### Clean up
```{r results=FALSE}
table(df1$from_hh, df1$from_indall, exclude=FALSE, deparse.level = 2)

# Filter to keep only those who appear in hhresp
df1 <- filter(df1, from_hh==1)

# drop the variables which indicate where the record came from
df1 <- select(df1, -from_hh, -from_indall)
```
## Females

```{r}
# i_indresp contains only adults aged 16 and above
dfind$flag_female <- dfind$i_sex_dv == 2

dfind_nfem    <- dfind %>%
                 group_by(i_hidp) %>%
                 count(flag_female, name="nfem") %>%
                 filter(flag_female) %>%
                 select(i_hidp, nfem) %>%
                 full_join(dfind, by="i_hidp") %>%
                 select(i_hidp, nfem) %>%
                 distinct(i_hidp, nfem) %>%
                 mutate_at(vars(nfem), ~replace(., is.na(.), 0))
```


### Merge data frames

Make sure to execute this only once
```{r}
df1$from_hh <- 1

dfind_nfem$from_indall <- 1

df1 <- full_join(dfind_nfem, df1, by="i_hidp")

```
### Clean up
```{r results=FALSE}

table(df1$from_hh, df1$from_indall, exclude=FALSE, deparse.level = 2)
# Filter to keep only those who appear in hhresp
df1 <- filter(df1, from_hh==1)

# drop the variables which indicate where the record came from
df1 <- select(df1, -from_hh, -from_indall)
```

## Education
```{r}
# Hiqual not missing

dfind_hiqual    <- dfind %>%
                 group_by(i_hidp) %>%
                 slice_min(order_by = i_hiqual_dv) %>%
                 rename(hh_hiqual = i_hiqual_dv) %>%
                 select(i_hidp, hh_hiqual) %>%
                 full_join(dfind, by="i_hidp") %>%
                 select(i_hidp, hh_hiqual) %>%
                 distinct(i_hidp, hh_hiqual) %>%
                 # Recode "missing hiqual" to 10, maybe also rename levels
                 mutate_at(vars(hh_hiqual), ~replace(., . == -8, 10)) %>%
                 mutate_at(vars(hh_hiqual), ~replace(., . == -9, 10))


```


### Merge data frames

Make sure to execute this only once
```{r}
df1$from_hh <- 1

dfind_hiqual$from_indall <- 1

df1 <- full_join(dfind_hiqual, df1, by="i_hidp")

```
### Clean up
```{r}

table(df1$from_hh, df1$from_indall, exclude=FALSE, deparse.level = 2)
# Filter to keep only those who appear in hhresp
df1 <- filter(df1, from_hh==1)

# drop the variables which indicate where the record came from
df1 <- select(df1, -from_hh, -from_indall)


```

Add a new label for the value 10 introduced above that correponds to "Missing qualification data"

```{r}
library("labelled")
# The 7th level of hh_hiqual corresponds to Missing data
val_label(df1$hh_hiqual, 10) <- "Missing `qualification`"


```

```{r}
sum(is.na(df1$hh_hiqual))
```


### Price data and conversion factors


Fuel price data and conversion factors are available from the Department of Energy and Climate Change (DECC), for details see [@buchs2013emits]. Gas prices for Northern Ireland (NI) are available from the Annual Transparency Report of NI's Utility Regulator (UREGNI). 


```{r message=FALSE}
library(readxl)
p_elec_paymentmethods <- read_excel("data/raw/p_elec_paymentmethods.xlsx")
p_gas_paymentmehods <- read_excel("data/raw/p_gas_paymentmehods.xlsx")
p_oil <- read_excel("data/raw/p_oil_prices.xlsx")

# Combine price tables into one dataframe
prices_df2 <- cbind(p_elec_paymentmethods,p_gas_paymentmehods,p_oil)
# Conversion factors
co2_df3 <- read_excel("data/raw/co2_df3.xlsx")
```

The DECC offers price data for credit, debit and prepayment. However USoc
reported several other payment methods. Hence we need a mapping from the methods
to the available price data.

```{r}
payment_mapping <- c("debit", "debit", "debit", "prepay", "prepay", "credit", "debit", "debit", "debit")
```

## Re-coding missing values 

We use the function `na_values` from the package `labelled` to re-code the values `97, -9, -8, -7, -2, -1` as 
user-defined missing values and to convert them to `NA`.

```{r}
for (i in 1:ncol(df1)) {
  na_values(df1[[i]]) <- c(97, -9, -8, -7, -2, -1)
}
```

The missing labels for the variable `i_psu` don't include `97`. Actually there 
are no missing values in `i_psu` at all. Since there are some households that 
have an `i_psu` value of `97` we have to remove `97` from the na-labels of that 
variable.

```{r}
na_values(df1$i_psu) <- c(-9, -8, -7, -2, -1)
```

Finally we can transform all user-defined missing values into NA

```{r}
df1 <- user_na_to_na(df1)
```

## Assesing some missing data

There are 66 households who did not report on the fuels they use.

```{r}
# The rowSums are either 0 or 4 which means either all are NA or none
df1 %>% select(i_fuelhave1,i_fuelhave2, i_fuelhave3, i_fuelhave4) %>% is.na  %>% rowSums %>% unique  # == c(0,4)

no_fueldat <- is.na(df1$i_fuelhave1)
```

When a household reports spending on gas and electricity separately they should be
considered to receive separate bills.
```{r}
gas_and_elec_xp <- !is.na(df1$i_xpgasy) & df1$i_xpgasy > 0  &
               !is.na(df1$i_xpelecy) & df1$i_xpelecy > 0

df1$i_fuelduel[gas_and_elec_xp] <- 2 # applies to 5 households



# Some checks
sum(is.na(df1$i_xpelecy) & is.na(df1$i_xpgasy) & df1$i_fuelduel == 2, na.rm=T) / sum(df1$i_fuelduel == 2, na.rm=T)

sum(is.na(df1$i_xpelecy) & is.na(df1$i_xpgasy))

df1$i_fuelduel %>% is.na %>% sum
```




## Imputing duel fuel gas to electricity ratio with nean imputation

Many  households (HHs) have a combined gas and electricity bill. They reported
only the combined expenditure and hence direct conversion of
their expenditures to CO2 consumption is not possible.
Since this is the case for many households we want to include them in our analysis nevertheless.

```{r}
sum(df1$i_fuelduel == 1, na.rm = T) # Dual Fuel
sum(df1$i_fuelduel == 2, na.rm = T) # Split Bill

```


In order to do this we compute the ratios of gas and electricity expenditure to
combined expenditure for HHs with split bills. The mean of this
ratio will be used to artificially split the dualfuel
bill into estimated direct expenditures for gas and electricity. 



```{r}
# Initialize columns for ratios
df1$gas_ratio <- as.numeric(rep(NA, nrow(df1)))
df1$elec_ratio <- as.numeric(rep(NA, nrow(df1)))
```

We can only compute the splitting ratio for people that have gas AND electricity.
Additionally we require those HHs to not use other fuels, as this will influence the ratio.
These households are removed later on anyhow because we can't reliably estimate their emissions
since we lack a conversion factor for other fuels.

```{r}
no_other_fuels <- !is.na(df1$i_fuelhave4) & df1$i_fuelhave4 == 0 


gas_or_elec_xp <- (!is.na(df1$i_xpgasy) & df1$i_xpgasy > 0  |
               !is.na(df1$i_xpelecy) & df1$i_xpelecy > 0 ) & no_other_fuels 


df1[gas_and_elec_xp,]$gas_ratio <- (df1[gas_and_elec_xp,]$i_xpgasy / 
                                (df1[gas_and_elec_xp,]$i_xpelecy + df1[gas_and_elec_xp,]$i_xpgasy))
df1[gas_and_elec_xp,]$elec_ratio <- 1 - df1[gas_and_elec_xp,]$gas_ratio
```

Further we differentiate between HHs with and without oil as additional energy source and 
construct two corresponding logical vectors for subsetting,

Since splitting the expenditure is only possible when expenditure was reported,
we need an additional vector.

```{r}
hh_no_oil <- !is.na(df1$i_fuelhave3) & df1$i_fuelhave3 == 0
hh_oilhav <- !is.na(df1$i_fuelhave3) & df1$i_fuelhave3 == 1
hh_duelxp <- !is.na(df1$i_xpduely) & df1$i_xpduely > 0
```

For HHs with duelfuel bill we derive the payment methods for gas and electricity
from the duelfuel variable.

```{r}
hh_dualfuel = !is.na(df1$i_fuelduel) & df1$i_fuelduel == 1
df1[hh_dualfuel,]$i_elecpay <- df1[hh_dualfuel,]$i_duelpay
df1[hh_dualfuel,]$i_gaspay <- df1[hh_dualfuel,]$i_duelpay 
```

We incorporate survey design weights to better account for bias in
the survey when estimating the splitting ratio.

```{r message=FALSE}
library("survey")

options(survey.lonely.psu = "adjust")

svy_df <- svydesign(id=~i_psu,
                    strata=~i_strata,
                    weights=~i_hhdenui_xw,
                    data=df1)
```
Total weight
```{r}
sum(df1$i_hhdenui_xw)
```

### Justification for mean imputation

Mean imputation is the most straightforward and easy to understand imputation strategy, however it has a number of disadvantages compared to other strategies. For example it will reduce the variance of the imputed variable. In our case the variance of the imputed variable before imputation is already small, due to the pronounced peak in the distribution of the electricity to gas ratio at approximately 0.5.


```{r}

pdf("plots/elec_ratio.pdf",10,7)
par(cex.lab=1.5, cex.axis=1.5, cex.main=1.5, cex.sub=1.5)
hist(df1[ gas_and_elec_xp & no_other_fuels,]$elec_ratio, 
     breaks=100,
     xlab = "Electricity to combined expenditure ratio",
     main = "")
dev.off()
# hist(df1[ gas_and_elec_xp & no_other_fuels,]$total_co2, breaks=100)
# hist(df1[ hh_duelxp & no_other_fuels,]$elec_ratio, breaks=100)
# hist(df1[ hh_duelxp & no_other_fuels,]$total_co2, breaks=100)
```


## Stochastic imputation

```{r}
library(mice) 
library(dplyr)

# It might be worthwhile to exclude hhs with incomplete oil expenditure
# sum((hh_duelxp | gas_and_elec_xp ) & no_other_fuels & !unknown_oil & !zero_oil , na.rm=T)
# AND to not only consider OILHAVE but also XPOILY in mice

# Impute dual fuel households by ratio of people who have elec and gas and
# reported their expendtiure. Exclude people with "other fuels".
dfimp <- df1[(hh_duelxp | gas_and_elec_xp ) & no_other_fuels,] %>% select(elec_ratio,i_fihhmnnet1_dv,i_hhsize, i_nkids_dv,i_nonepar_dv,i_ncouple_dv,i_npensioner,i_nue_dv,nfem,nhcond,i_hsbeds,i_hsrooms,i_hsownd,hh_hiqual,i_gor_dv,i_urban_dv, i_fborn, i_fuelhave3,i_heatch)
dfimp <- dfimp[,2:ncol(dfimp)]


dfimp[1:12] <- zap_labels(dfimp[1:12])
dfimp[13:ncol(dfimp)] <- droplevels( as_factor(dfimp[13:ncol(dfimp)]))

# This is one way to specify that only elec_ratio should be imputed
# whereMatrix <- dfimp
# whereMatrix$elec_ratio <- is.na(whereMatrix$elec_ratio)
# whereMatrix[,2:ncol(whereMatrix)] <- FALSE

# Drop predictors with missing data (the most important one (oil) has no missings)
dfimp <- cbind.data.frame(dfimp[1], dfimp[,colSums(is.na(dfimp)) == 0])



imp <- mice(dfimp, method = "pmm", m=1)#, where = whereMatrix)


# hist(df1$elec_ratio, breaks=100)      
# hist(imp$imp$elec_ratio[,1], breaks=100)      


var(imp$imp$elec_ratio[,1], na.rm=T)
var(imp$data$elec_ratio, na.rm=T)

mean(imp$imp$elec_ratio[,1], na.rm=T)
mean(imp$data$elec_ratio, na.rm=T)
```


```{r}
df1[(hh_duelxp | gas_and_elec_xp ) & no_other_fuels,]$elec_ratio <- complete(imp)$elec_ratio
df1$gas_ratio <- 1 - df1$elec_ratio
```


Finally we assign the weighted and differentiated mean splitting ratios, to the HHs 
with duel fuel bill and use that ratio to impute the gas and electricity expenditure.
For the stochastic imputation we use the ratio from mice directly.

```{r}
# Only for mean imputation strategy
# df1[hh_duelxp & hh_no_oil,]$elec_ratio <- svymean(~elec_ratio, subset(svy_df, gas_and_elec_xp & hh_no_oil & no_other_fuels))[1]
# df1[hh_duelxp & hh_no_oil,]$gas_ratio  <- svymean(~gas_ratio , subset(svy_df, gas_and_elec_xp & hh_no_oil & no_other_fuels))[1]
# df1[hh_duelxp & hh_oilhav& no_other_fuels,]$elec_ratio <- svymean(~elec_ratio, subset(svy_df, gas_and_elec_xp & hh_oilhav))[1]
# df1[hh_duelxp & hh_oilhav& no_other_fuels,]$gas_ratio  <- svymean(~gas_ratio , subset(svy_df, gas_and_elec_xp & hh_oilhav))[1]


# For mean and stochastic imputation
df1[hh_duelxp,]$i_xpelecy  <- df1[hh_duelxp,]$i_xpduely * df1[hh_duelxp,]$elec_ratio
df1[hh_duelxp,]$i_xpgasy   <- df1[hh_duelxp,]$i_xpduely * df1[hh_duelxp,]$gas_ratio
```

The total weight of all households for which we estimate the ratio is
```{r}
svytotal(~I(i_hhdenui_xw > -1), subset(svy_df, hh_duelxp & no_other_fuels))
```
... and for HH for which the ratio is available in the data.
```{r}
svytotal(~I(i_hhdenui_xw > -1), subset(svy_df, gas_and_elec_xp & no_other_fuels))
```
The electricity / gas ratio for HH without oil is
```{r}



mean(df1[hh_duelxp & no_other_fuels,]$elec_ratio / df1[hh_duelxp & no_other_fuels,]$gas_ratio)
mean(df1[gas_and_elec_xp& no_other_fuels & hh_no_oil,]$elec_ratio / df1[gas_and_elec_xp& no_other_fuels & hh_no_oil,]$gas_ratio)
```

```{r}
df1[hh_duelxp & hh_oilhav,]$elec_ratio / df1[hh_duelxp & hh_oilhav,]$gas_ratio
```
The total mean of elec to gas for the dataset is
```{r}
r = svymean(~elec_ratio, svy_df, na.rm =T)
svymean(~i_xpelecy, svy_df, na.rm =T) / svymean(~i_xpgasy, svy_df, na.rm =T)
```



## Computation of consumption 


We start by adding variables that contain the exact prices each HH has to pay
for the different energy sources. These variables
are initialized as NA to later account for missing data.

```{r}

df1$elec_prize <- as.numeric(rep(NA, nrow(df1)))
df1$gas_prize <- as.numeric(rep(NA, nrow(df1)))
df1$oil_prize <- as.numeric(rep(NA, nrow(df1)))
```

To compute the prizes we need information on survey year, region and payment method (only for gas and electricity).
Since some HH miss that information we need to select suitable subsets first

```{r}
not_missing_gor_or_yr = !is.na(df1$i_intdatey) & !is.na(df1$i_gor_dv)
elecpay_method_available = !is.na(df1$i_elecpay)
gaspay_method_available  = !is.na(df1$i_gaspay)
```

Adding the prize variable only makes sense if the expenditure for the corresponding
fuel has been recorded. 

```{r}
elec_expenditure_available = !is.na(df1$i_xpelecy) 
gas_expenditure_available  = !is.na(df1$i_xpgasy)
oil_expenditure_available  = !is.na(df1$i_xpoily)
```

This gives us the subsets "good" people for which the prizes can be straightforwardly computed

```{r}
good_people_elec = not_missing_gor_or_yr & elec_expenditure_available & elecpay_method_available
good_people_gas  = not_missing_gor_or_yr & gas_expenditure_available & gaspay_method_available
good_people_oil  = not_missing_gor_or_yr & oil_expenditure_available
```

To look up the prizes in the prize data frame we define three functions, one for each source of energy.
These functions will be applied to the HH dataframe in order to add the price data for each row.

```{r}

getElecPrize <- function(row){
  year            <- toString(row[["i_intdatey"]] - 1) # with - 1 
  region          <- row[["i_gor_dv"]]
  payment_number  <- row[["i_elecpay"]]
  payment_method  <- payment_mapping[payment_number]
  column_name_df2 <- paste(year, "elec", payment_method, sep="_")
  price_in_pound  <- prices_df2[region, column_name_df2] / 100
  return(price_in_pound)
}

getGasPrize <- function(row){
  year            <- toString(row[["i_intdatey"]] - 1) # with - 1 
  region          <- row[["i_gor_dv"]]
  payment_number  <- row[["i_gaspay"]]
  payment_method  <- payment_mapping[payment_number]
  column_name_df2 <- paste(year, "gas", payment_method, sep="_")
  price_in_pound  <- prices_df2[region, column_name_df2] / 100
  return(price_in_pound)
}

getOilPrize <- function(row){
  year            <- toString(row[["i_intdatey"]] - 1) # with - 1 
  region          <- row[["i_gor_dv"]]
  column_name_df2 <- paste(year, "oil","prize", sep="_")
  price_in_pound_per_litre  <- prices_df2[region, column_name_df2] / 100
  # We don't need a conversion factor here if we have one for litres -> co2
  price_in_pound <- price_in_pound_per_litre
  return(price_in_pound)
}


df1[good_people_elec, "elec_prize"] <- apply(df1[good_people_elec,], 1, getElecPrize)
df1[good_people_gas, "gas_prize"]   <- apply(df1[good_people_gas,], 1, getGasPrize)
df1[good_people_oil, "oil_prize"]   <- apply(df1[good_people_oil,], 1, getOilPrize)
```


## Imputation of missing price data

Unfortunately there are some people for which expenditure data is available
but either government region or year or payment method is missing. 
For those cases we will perform data imputation with weighted (categorical) means.

We chose mean imputation since it is fairly simple to implement and since
we only impute the price that an household has to pay. The main information affecting the CO2 emissions is still contained in the expenditure variabels.

We start out by assessing how many prices have to be imputated
```{r}
sum(elec_expenditure_available & !elecpay_method_available) 
sum(elec_expenditure_available & !not_missing_gor_or_yr)
sum(gas_expenditure_available & !gaspay_method_available)
sum(gas_expenditure_available & !not_missing_gor_or_yr)
sum(oil_expenditure_available & !not_missing_gor_or_yr)
```

We define two functions that impute the prize with the mean of all households that
have the same year, region or payment method respectively. When computing the means
we take the survey weights of the corresponding subset into account

Since imputation in the way implemented below is an iterative process,
working line by line, the mean ElecPrize will change after a row has been imputed.

Therefor we make a copy of df1 before imputation, that serves as the baseline on which the means are computed.

The functions defined below are implemented in a straightforward but computationally inefficient manner.
The surveymean is recomputed for each combination of missing values when it would be more efficient
to cache already computed values and reuse them. However these functions have to be run only once.

```{r}
copy_for_mean_computation_df1 = data.frame(df1)
# we use the survey package to equip that copy with weights
svy_cp <- svydesign(id=~i_psu,
                     strata=~i_strata,
                     weights=~i_hhdenui_xw,
                     data=copy_for_mean_computation_df1)

imputeElecPrize <- function(row){
  # If payment_method is missing
  if (is.na(row[["i_elecpay"]])) {
    year <- row[["i_intdatey"]]
    region <- row[["i_gor_dv"]]
    # doing it like this for every row is computationally inefficient but well...
    svy_sub <-subset(svy_cp, good_people_elec & df1$i_intdatey == year & df1$i_gor_dv == region)
    price_in_pound <- svymean(~elec_prize, svy_sub)[[1]]
    return(price_in_pound)
  }
  # If year is unknown
  if (is.na(row[["i_intdatey"]])) {
    region <- row[["i_gor_dv"]]
    payment_number <- row[["i_elecpay"]]
    svy_sub <-subset(svy_cp, good_people_elec & df1$i_elecpay == payment_number & df1$i_gor_dv == region)
    price_in_pound <- svymean(~elec_prize, svy_sub)[[1]]
    return(price_in_pound)
  }
  # If region is unknown
  if (is.na(row[["i_gor_dv"]])) {
    year <- row[["i_intdatey"]]
    payment_number <- row[["i_elecpay"]]
    svy_sub <-subset(svy_cp, good_people_elec & df1$i_elecpay == payment_number & df1$i_intdatey == year)
    price_in_pound <- svymean(~elec_prize, svy_sub)[[1]]
    return(price_in_pound)
  }
}

imputeGasPrize <- function(row){
  # If payment_method is missing
  # Unfortunately there is one HH with unknown gas_pay AND unknown region
  if (is.na(row[["i_gaspay"]]) & is.na(row[["i_gor_dv"]])) {
    year <- row[["i_intdatey"]]
    svy_sub <-subset(svy_cp, good_people_gas & df1$i_intdatey == year)
    price_in_pound <- svymean(~gas_prize, svy_sub)[[1]]
    return(price_in_pound)
  } else if (is.na(row[["i_gaspay"]])) {
    # If only gaspay is unknown
    year <- row[["i_intdatey"]]
    region <- row[["i_gor_dv"]]
    svy_sub <-subset(svy_cp, good_people_gas & df1$i_intdatey == year & df1$i_gor_dv == region)
    price_in_pound <- svymean(~gas_prize, svy_sub)[[1]]
    return(price_in_pound)
  } else if (is.na(row[["i_gor_dv"]])) {
    # If only region is unknown
    year <- row[["i_intdatey"]]
    payment_number <- row[["i_gaspay"]]
    svy_sub <-subset(svy_cp, good_people_gas & df1$i_gaspay == payment_number & df1$i_intdatey == year)
    price_in_pound <- svymean(~gas_prize, svy_sub)[[1]]
    return(price_in_pound)
  }
  # If year is unknown
  if (is.na(row[["i_intdatey"]])) {
    region <- row[["i_gor_dv"]]
    payment_number <- row[["i_gaspay"]]
    svy_sub <-subset(svy_cp, good_people_gas & df1$i_gaspay == payment_number & df1$i_gor_dv == region)
    price_in_pound <- svymean(~gas_prize, svy_sub)[[1]]
    return(price_in_pound)
  }
}
```

The functions for electricty and gas price imputation can be applied straightforwardly.

```{r}
# For electricity
df1[elec_expenditure_available & !good_people_elec, "elec_prize"] <- apply(df1[elec_expenditure_available & !good_people_elec,], 1, imputeElecPrize)

# For gas

df1[gas_expenditure_available & !good_people_gas, "gas_prize"] <- apply(df1[gas_expenditure_available & !good_people_gas,], 1, imputeGasPrize)
```

There is just one HH with missing incomplete oil prize data which we impute individually (region missing).

```{r}
oil_impute_year <- df1[oil_expenditure_available & !not_missing_gor_or_yr,]$i_intdatey
oil_svy_sub <- subset(svy_cp, good_people_oil & df1$i_intdatey == oil_impute_year)
oil_impute_price_in_pound <- svymean(~oil_prize, oil_svy_sub)[[1]]
df1[oil_expenditure_available & !not_missing_gor_or_yr,]$oil_prize <- oil_impute_price_in_pound
```

With all  prize information stored in the data frame calculating the fuel consumption
is easy.

```{r}
df1$elec_consumption <- df1$i_xpelecy / df1$elec_prize
df1$gas_consumption  <- df1$i_xpgasy / df1$gas_prize
df1$oil_consumption  <- df1$i_xpoily / df1$oil_prize
```


## Computation of CO2 emissions


The process is in principle very similar to the computation of consumption data

While the conversion factors for oil and gas are (approximately) the same for
different years, the factor for electricity differs by almost 50%. 

Since we are not primarily interested in absolute energy consumption emissions,
but in relative emissions and their dependence on socio-economic factors,
we have to ensure comparability across multiple years.
To achieve that we compute weighted mean conversion factors for electricty.

To do this we sum the relative survey weight of HHs responding in each year (stored in the 
variable `i_hhdenui_xw`) and multiply with conversion factor for that year. 

The original factors are stored in co2_df3.

```{r}
mean_elec_co2_factor= (sum(df1[df1$i_intdatey == 2017,]$i_hhdenui_xw, na.rm = T) * 0.44932 + 
                            sum(df1[df1$i_intdatey == 2018,]$i_hhdenui_xw, na.rm = T) * 0.38443 +
                            sum(df1[df1$i_intdatey == 2019,]$i_hhdenui_xw, na.rm = T) * 0.3072) /
                            sum(df1[df1$i_intdatey > 2016,]$i_hhdenui_xw, na.rm = T)
mean_gas_co2_factor  = 0.184
mean_oil_co2_factor  = 2.53
```

CO2 emissions are then computed as fuel consumption multiplied by the corresponding factor.

```{r}
df1$elec_co2 <- df1$elec_consumption * mean_elec_co2_factor
df1$gas_co2  <- df1$gas_consumption * mean_gas_co2_factor
df1$oil_co2  <- df1$oil_consumption * mean_oil_co2_factor
```

Finally we sum the emission due to each energy source to arrive at an estimate of
the total annual household CO2 emission due to energy consumption.

The following code does that without producing `NA` when at least one type of emission is given.

```{r}

getTotalCo2 <- function(row){
  total <- 0.
  if (!is.na(row[["elec_co2"]])) { total <- total + row[["elec_co2"]] }
  if (!is.na(row[["gas_co2"]]))  { total <- total + row[["gas_co2"]] }
  if (!is.na(row[["oil_co2"]]))  { total <- total + row[["oil_co2"]] }
  if (is.na(row[["elec_co2"]]) & is.na(row[["gas_co2"]]) & is.na(row[["oil_co2"]]))
                                 { total <- NA }
  return(total)
}

df1$total_co2 <- apply(df1, 1, getTotalCo2)
```



Finally we can check that all households who reported some kind of expenditure on
electricity, gas or oil have been considered

```{r}
sum(df1$i_xpgasy > 0 | df1$i_xpelecy > 0 | 
    df1$i_xpoily > 0 | df1$i_xpduely > 0 & no_other_fuels, na.rm=T) == sum(df1$total_co2 > 0, na.rm=T)


```



##  Contrasts for linear regression

When using factor variables in linear regressions a baseline has to be chosen
against which the different levels can be compared. The coefficients of the model
encode relative changes, with respect to that baseline. By default the first 
factor level is used as baseline. Sometimes that leads to hard to interpret 
coefficients, e.g. when the baseline is an atypical or little interesting case.
In this cases a contrast may be used to encode another baseline.

The following code changes the baseline of the variables for house ownership and
household composition to be equal to the mean of all levels, a so-called
**sum contrast**.

```{r message=FALSE}
library("labelled")
for (rowname in c("i_hsownd", "i_hhtype_dv", "hh_hiqual")) {
  df1[[rowname]] <- to_factor(df1[[rowname]],  drop_unused_labels = T)
  contrasts(df1[[rowname]]) <- contr.sum(length(levels(df1[[rowname]])))
}
```

Since independent variables in a regression model can not be perfectly collinear
one of the factor levels has to be dropped nevertheless. It's coefficient can be
obtained by a linear combination of all other coefficients. The above code drops
the last levels. However for government region the last level is of particular
interest since it corresponds to Northern Ireland (NI), and later we will see that,
NI's coefficient  strongly deviates from the mean. Contrarily, the first level to 
the North East, whose regression coefficent is close to the mean. Hence, we choose
a sum contrast that does not drop NI but instead North East:

```{r}
for (rowname in c("i_gor_dv")) {
  df1[[rowname]] <- to_factor(df1[[rowname]],  drop_unused_labels = T)
  nrow_contr <- length(levels(df1[[rowname]]))
  contrasts(df1[[rowname]]) <- rbind(rep(-1,nrow_contr-1), diag(nrow_contr-1))
}
```
Now here a small caveat applys. This procedure lower the integer value associated
to every level by 1, such that NI now correspond to level 11 instead of 12.

## Subset selection

The dataset still contains many households (HHs) for which no reliable CO2 emissions
can be estimated. Therefor we have to select a suitable subset of `good_people` that
fulfill all requirements and will be used for further analysis. We will create logical
vectors and intersect them to arrive at the subset vector of `good_people`.

We start be excluding HH with "other fuels" since for those no conversion factors
exists and hence we would underestimate their CO2 emissions. These might be included
in further analyses since the total expenditure of other fuels is small compared to those
of gas, oil and electricity.

```{r}
other_fuels = !is.na(df1$i_fuelhave4) & df1$i_fuelhave4 == 1
```

We further exclude HH that have a certain fuel but don't specify how much they spent on it.
They might have their own solar panels or an oil well, which would imply
vastly different CO2 emissions.
```{r}
zero_elec = !is.na(df1$i_fuelhave1) & df1$i_fuelhave1 == 1 &
               !is.na(df1$i_xpelecy) & df1$i_xpelecy == 0

zero_gas = !is.na(df1$i_fuelhave2) & df1$i_fuelhave2 == 1 &
              !is.na(df1$i_xpgasy) & df1$i_xpgasy == 0

zero_oil = !is.na(df1$i_fuelhave3) & df1$i_fuelhave3 == 1 &
              !is.na(df1$i_xpoily) & df1$i_xpoily == 0


unknown_elec = !is.na(df1$i_fuelhave1) & df1$i_fuelhave1 == 1 &
               is.na(df1$i_xpelecy) 

unknown_gas = !is.na(df1$i_fuelhave2) & df1$i_fuelhave2 == 1 &
              is.na(df1$i_xpgasy)

unknown_oil = !is.na(df1$i_fuelhave3) & df1$i_fuelhave3 == 1 &
              is.na(df1$i_xpoily)


```

Finally there are some HHs for which we couldn't estimate CO2 emission due to lacking data
```{r}
unknown_co2 = is.na(df1$total_co2)
```

HHs that fall into none of these categories are "good".

```{r}
known_people = !(other_fuels | zero_elec | zero_gas | zero_oil| unknown_elec | unknown_gas | unknown_oil | unknown_co2)
```

Let's have a look at how much HHs belong into each category.

```{r}
sum(other_fuels)
sum(unknown_elec)
sum(unknown_gas)
sum(unknown_oil)
sum(zero_elec)
sum(zero_gas)
sum(zero_oil)
sum(unknown_co2)
sum(!known_people)
```

## Regularization

Some households have especially high or low carbon emissions. Most of those
are likely to be results of data collection errors or incomplete responses. 
We regularize the data by discarding the samples in the top and bottom half
percentile of total carbon emissions.

```{r}
 
bottom_half_percentile = known_people & df1$total_co2 < quantile(df1[known_people,]$total_co2, 0.005, na.rm=T)
top_half_percentile = known_people & df1$total_co2 > quantile(df1[known_people,]$total_co2, 0.995, na.rm=T)

sum(known_people) / sum(bottom_half_percentile | top_half_percentile) # sanity check

good_people = known_people & !(bottom_half_percentile | top_half_percentile)

df1$good <- good_people

dfgood = df1[good_people,]
```

The final subset used for our analysis contains `r sum(good_people)` of `r nrow(df1)` households.

## Adding survey weights

Finally we can add weights and start the analysis.

We define two different data sets, svy_all and svy_good
- svy_all: contains all data, and is used e.g. for income analysis
- svy_good: contains all **good** HHs, we use it for CO2 emissions analsyis

```{r}
svy_all <- svydesign(id=~i_psu,
                     strata=~i_strata,
                     weights=~i_hhdenui_xw,
                     data=df1)
```


## Income Deciles

With the weights attached to the data we can compute weighted deciles of disposable 
and equivalised household net incomes. First compute the deciles...

```{r}
hh_income_deciles <- svyquantile(~i_fihhmnnet1_dv, svy_all, seq(from=0.1, to=0.9, by=0.1))
eq_income_deciles <- svyquantile(~eq_income, svy_all, seq(from=0.1, to=0.9, by=0.1))
```

Then update the `survey` object with a new column

```{r}
svy_all <- update(svy_all, 
                  hh_income_decile = factor(1 + findInterval(i_fihhmnnet1_dv, hh_income_deciles)))
svy_all <- update(svy_all, 
                  eq_income_decile = factor(1 + findInterval(eq_income, eq_income_deciles)))
```

As expected the income increases from decile to decile.

```{r}
svyby(~eq_income, ~eq_income_decile, svy_all, svymean)
```

Now we create the `svy_good` object which contains the HHs suitable for CO2 analysis.
It inherits the variables for the income deciles from the `svy_all` object.

```{r}
svy_good <- subset(svy_all, good_people)
```

We add a new column for per capita CO2 emissions

```{r}

svy_good <- update(svy_good, per_capita_co2 = total_co2 / i_hhsize)
```



## Preparing emission quartile analysis

Quartiles w.r.t. total_co2:
```{r}
total_co2_quartiles <- svyquantile(~total_co2, svy_good, seq(from=0.25, to=0.75, by=0.25))

svy_good <- update(svy_good, 
                   total_co2_quartile = factor(1 + findInterval(total_co2, total_co2_quartiles)))
```
Quartiles w.r.t. per_capita_co2:
```{r}
per_capita_co2_quartiles <- svyquantile(~per_capita_co2, svy_good, seq(from=0.25, to=0.75, by=0.25))

svy_good <- update(svy_good, 
                   per_capita_co2_quartile = factor(1 + findInterval(per_capita_co2, per_capita_co2_quartiles)))
```


## Limitations 

### Imputation of households with dual fuels

Here are some statistics that differ between single and dual fuellers. The ratio we derive from the single bill ratio and use to split the dual bills is however in good accordance with data from official sources.

```{r}
rn = c("Separate Bills", "Dual Fuel")
lenr = length(rn)
cn = c("eq. income", "# pax", "# kids", "# bedrooms", "% oil heating")
lenc = length(cn)


S_cont <- svyby(~ eq_income / i_hhsize / i_nkids_dv / i_hsbeds / i_fuelhave3, ~I(i_fuelduel==1), svy_all, svymean, na.rm = T)
# Initialize table
tab <- sprintf("%i (%i)", round(S_cont[,2]), round(S_cont[,2+lenc]))
# Add rows
for (i in 3:5){
  tab <- cbind(tab, sprintf("%.2f (%.2f)", S_cont[,i], S_cont[,i+lenc]))
}
i=i+1
  tab <- cbind(tab, sprintf("%.2f (%.2f)", 100 * S_cont[,i], 100 *  S_cont[,i+lenc]))

ptab <- matrix(tab,lenr, length(cn),dimnames=list(rn,cn))
ptab
```
Latex table
```{r eval=FALSE}
library("xtable")
xtable(ptab,
       caption = "Comparison of relevant characteristics of households with and without dual fuel deal.\\label{tab:comp_dualfuel}")
```

### Other fuels
We did not consider emissions from other fuels. The expenditure on those fuels is comparatively small. Nevertheless, it might be interesting to study those in future work.
```{r}
sum(df1$i_xpsfly, na.rm=T) / 
  (sum(df1$i_xpelecy, na.rm=T) + sum(df1$i_xpoily, na.rm=T) + sum(df1$i_xpgasy, na.rm=T))
```


### Subset bias

Our process of selecting the subset of good people introduces bias into the dataset.

Here is a table showing differences betweens some of the regressors for the used subset and the  dropped values.

```{r}
rn = c("Dropout", "Used Subset", "Complete Survey")
lenr = length(rn)
cn = c("eq. income", "# pax",  "# bedrooms", "% oil heating")
lenc = length(cn)

rowform = ~eq_income / i_hhsize / i_hsbeds / i_fuelhave3 

S_cont <- svyby(rowform, ~good, svy_all, svymean, na.rm = T)
S_ext <- svyby(rowform, ~I(1), svy_all, svymean, na.rm = T)
names(S_ext)[1] <- "good"
S_cont <- rbind(S_cont,S_ext)
# Initialize table
tab <- sprintf("%i (%i)", round(S_cont[,2]), round(S_cont[,2+lenc]))
# Add rows
for (i in 3:4){
  tab <- cbind(tab, sprintf("%.2f (%.2f)", S_cont[,i], S_cont[,i+lenc]))
}
i=i+1
tab <- cbind(tab, sprintf("%.2f (%.2f)", 100 * S_cont[,i], 100 * S_cont[,i+lenc]))

ptab <- matrix(tab,lenr, length(cn),dimnames=list(rn,cn))
ptab
```
Latexify
```{r eval=FALSE}
xtable(ptab,
       caption = "Mean (standard deviation) of important variables for the subset of dropped HHs, the subset of used HHs and the whole dataset\\label{tab:comp_dropout}.")
```


Complete comparison table for reference

```{r}
rowform = ~eq_income / 
          total_co2 /
          log(i_fihhmnnet1_dv + 961) /
          i_hsownd /
          # Education
          hh_hiqual /
          # Region
          i_gor_dv /
          i_urban_dv /
          # Household Size
          i_hhsize / 
          I(i_hhsize == 1) /
          # Household Composition
          i_nkids_dv /
          i_nonepar_dv / 
          i_ncouple_dv / 
          i_npensioner /
          i_nue_dv /
          nfem / 
          nhcond /
          i_fborn /
          # Accommodation 
          i_hsbeds /
          i_hsrooms / 
          as.factor(i_fuelhave3) /
          i_heatch

S_cont <- svyby(rowform, ~good, svy_all, svymean, na.rm = T)
S_ext <-svyby(rowform, ~I(1), svy_all, svymean, na.rm = T)
names(S_ext)[1] <- "good"
S_cont <- rbind(S_cont,S_ext)
S_cont
```



Total weight of the used subset

```{r}
sum(good_people)
svytotal(~good, svy_good, na.rm=T)
svytotal(~I(i_fuelduel==1), svy_good, na.rm=T)
```

## Exporting the preprocessed data

This concludes the preprocessing. We export the data for later use in other scripts.

``` {r}
write_sav(df1, "data/i_hhresp_preprocessed_mice.sav")
write_sav(dfgood, "data/dfgood_mice.sav")
saveRDS(svy_all, "data/svy_all_mice.rds")
saveRDS(svy_good, "data/svy_good_mice.rds")
```

# References